---
title: "Homework 2"
author: "JR Chansakul"
date: 2020-09-21
output: 
  html_document:
    toc: true
---

```{r}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset. 

```{r}
trashwheel_df =
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_ball = round(sports_balls),
    sports_ball = as.integer(sports_balls)
  )
```

This dataset contains information for the Mr. Trashwell trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash 

Read precipitation data!

```{r}
precip_2018 =
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>%  
  janitor::clean_names() %>% 
  drop_na(month) %>% 
    mutate(year = 2018) %>% 
    relocate(year)
  
precip_2017 =
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>%  
  janitor::clean_names() %>% 
  drop_na(month) %>% 
    mutate(year = 2017) %>% 
    relocate(year)

```

Now combine annual precipitation 

```{r}
month_df =
    tibble(
      month_num = 1:12
    )

precip_df = 
  bind_rows(precip_2018, precip_2017)
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.



# Problem 2

```{r NYC_transit_data}
NYC_transit_df =
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
    janitor::clean_names() %>% 
    select(line:ada, -staffing, -staff_hours, -exit_only)
     
    
  
    
```

